Squint: Lossy Hierarchical Compression on Symbol Streams

### 2012-12-06 tex/paper.tex
Began filling in results section, collected some final results

The previous algorithm updates the grammar sequentially, since each
variable replacement can have an effect on large parts of the grammar,
we will have to go through comparing all pairs again after the change.
An alternative algorithm would be to try to replace similar variables in
parallel. This would avoid many of the repeat pair-comparisons done in
\emph{Similarity}. In other words, the algorithm below makes puts 
similar variables
into equivalence class clusters, and then replaces them with a representative
all at once each iteration.

@@

\section{Results}
\subsection{Illustrative Examples}
Qualitative, show off diagrams and detailed metrics for small examples
\subsection{Similarity vs Clustering}
Fixed compression ratio + fidelity comparison for different ratios
Show the runtime difference
\subsection{Grammar Reduction Impact}
\subsection{Larger Tests}

\section{Conclusions}
\subsection{Ratios and Fidelity}
\subsection{Future Work}

>>>>

The previous algorithm updates the grammar sequentially, since each
variable replacement can have an effect on large parts of the grammar,
we will have to go through comparing all pairs again after the change.
An alternative algorithm would be to try to replace similar variables in
parallel. This would avoid many of the repeat pair-comparisons done in
\emph{Similarity}. In other words, the algorithm below makes puts 
similar variables
into equivalence class clusters, and then replaces each variable
in a cluster them with a representative
all at once each iteration.

@@

\section{Results}
\subsection{Illustrative Examples}
It is useful to look at how the algorithm functions on a small piece of
data with both plenty of hierarchical structure, and also some messiness
that most lossless compression schemes would not deal well with. Take
the string below, similar to what one might come across in a math
puzzle book.

\begin{verbatim}
123,124,123 + 321,323,321 = ???
124,123,123 + 321,3231,321 = ????
--
124,123,123 + 321,3231,321 = ????
123,124,123 + 321,323,321 = ????
\end{verbatim}

Disregarding its meaning, it consists of four very similar equations of the
form $a + b = ?$, and on a smaller scale each term consists of
3 repetitions of approximately the same string of either $123$ or $321$.

Sequitur yields a relatively complicated inferred grammar with 13 rules and
a size of 51. Both the Similarity and Clustering Algorithms with $\epsilon = .4$
, yield the following grammar after reduction. It has 5 rules and is almost
half the size, at 30.

\begin{verbatim}
~[*] => ~[BC]--~[BC]?

~[AZ] => ~[A]3~[A]~[A] + ~[C]~[C]3~[C]1 = ???
~[A] => 12
~[BC] => ~[AZ]~[AZ]
~[C] => 32
-----------------------
1231212 + 32323321 = ???1231212 + 32323321 = ???--(cont.)
1231212 + 32323321 = ???1231212 + 32323321 = ????
\end{verbatim}

The lossifier algorithm reduced the string to a pair of
pairs of the same equations, where each equation consists of the
sum of triples of "12" or "32". Smaller discrepencies
in the numbers involved were dropped, while the filler
commas and linebreaks also ended
up being lost. Note that the structure described is easily
observable from the grammar generated.

\subsection{Similarity vs Clustering}
For most of our benchmark tests, we focused on two extreme kinds of data:
one with large amounts of repetition at all levels, and one with a word
and sentence level structure but with little obvious large-scale repetition.
One set of test data was constructed by repeating the same letter `a' n-times
but then adding some
texture to the data by changing 1/30 of the characters uniformly
to `a'. Call these test-cases \emph{rep-n}
Another set of test data was constructed from the first n characters of the
book of Genesis from the KJV. Call these test-cases \emph{gen-n}

For our first suite of tests, we compare the effectiveness of the
\emph{Similarity} and \emph{Cluster} algorithms. We might expect
Similarity to achieve better quality at the same compression ratio
since it updates the state of the grammar between each variable
replacement, on the other hand since it must update the grammar
it has the potential to run much more slowly than Cluster.

We don't report the running times for just \Sequitur since its linear
running time is negligible compared to our algorithms for all but the
smallest input data.

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & 0.006 & 18/34    & 0.713 & 0.002 & 20/34    & 0.535\\
rep-400   & 0.028 & 45/76    & 0.269 & 0.009 & 41/76    & 0.743\\
rep-1600  & 0.355 & 103/194  & 0.873 & 0.042 & 103/194  & 0.804\\
rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
\end{tabular}
\caption{$\epsilon = .4$}
\end{table}

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & 0.002 & 30/34    & 0.040 & 0.002 & 30/34    & 0.040\\
rep-400   & 0.019 & 60/76    & 0.035 & 0.009 & 60/76    & 0.035\\
rep-1600  & 0.350 & 123/194  & 0.355 & 0.055 & 123/194  & 0.206\\
rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
\end{tabular}
\caption{$\epsilon = .2$}
\end{table}

TODO: Fill in results for $\epsilon = .1$

TODO: Runtime graphs and regression

Qualitatively the output strings generally are strings of a's with interspersed
`*' as expected, but the length of the final output varies widely.

Things to mention: Compression ratios and fidelity comparable for similarity
and cluster. Cluster has much better growth rate. Smaller epsilon provide
much better value/cost ratio as the input files get bigger for highly
repetitive files.

Thus, we see that similarity provides no consistent advantage over cluster
and has much less scalable performance, with an estimated running time of 
BLAH. For the rest of the paper we will only test Cluster.

\subsection{Grammar Reduction Impact}
benchmark two extreme test cases again, with and without reducer
see how much reducer helps (it shouldn't help too much)
notice how slow it is

\subsection{Larger Tests}
Final compression ratios on benchmark data
ASCII ART =p
Our Code

\section{Conclusions}
\subsection{Ratios and Fidelity}
\subsection{Future Work}
> Better ways to set epsilon to maximize value / cost for different size and types
> Better ways to quantify preservation of hierarchies
> How to find near `self-similarities' that aren't captured by a rule
> Iterating lossify and reduce in a cycle
